Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-01-12 response-to-reviews.Rmd

*[issue #92](https://github.com/EcoClimLab/vertical-thermal-review/issues/92)*

In the first part of Section 5.1, the authors conclude this paragraph stating the 'dominant role of vertical profiles in microclimate in shaping tree growth rates'. However, I believe the text has confused the very large effect of light with other aspects of forest microclimate, and they very much need to clarify this. If they believe other aspects of microclimate apart from light are drivers, then I believe this would be a minority opinion in the forest biology community. If they wish to speculate about other microclimate factors that cannot be disentangled from light, I suggest that they clearly indicate their opinion and speculation. It is perhaps an area for future research to disentangle light microclimate from aspects of vertical microclimate.

**We did not intend to argue that light was not the dominant factor, but the wording was misleading. This sentence has been revised to read, "This points to a dominant role of vertical profiles in the biophysical environment (Fig. 2) in shaping tree growth rates within forests."**
*[issue #91](https://github.com/EcoClimLab/vertical-thermal-review/issues/91)*

### 2022-01-12 response-to-reviews.Rmd

**We did not intend to argue that light was not the dominant factor, but the wording was misleading. This sentence has been revised to read, "This points to a dominant role of vertical profiles in the biophysical environment (Fig. 2) in shaping tree growth rates within forests."**

I am uncomfortable reading some of the major conclusions of the work here, such as speculation that large canopy trees are most vulnerable to warming when water is limited, but understory trees may be more vulnerable [to warming] when well-watered. This is not consistent with my knowledge of 'preponderance of available data'. Perhaps the authors implying that understory trees will succumb to warming earlier in a heat wave than large canopy trees? Is this based on the T50 evidence (l. 542-548) which is from 2 studies, or from one tree-ring study (l. 674-676), or both? An intriguing thought, but these studies and the corresponding author’s study have weaknesses and no clean experiment on this has been set up and done. I recommend that the more cautious language on l. 676-678 should in fact be repeated in the conclusions relevant to this point. It’s less stimulating and sober, but true, that it’s difficult to say so with strength of evidence and remains for further testing with good experimental design. An objective of such a review can stimulate further and better research.

**We agree that more caution is warranted on comparing canopy vs understory responses to warming, particularly because understory responses are highly uncertain. We have reworded relevant text as follows:**

### 2022-01-13 vertical-thermal-review.Rmd
fix issue #91
commiting to the change

**Authors:** 
Nidhi Vinod^1,2^,
Martijn Slot^3^,
Ian McGregor^4^,
Elsa M. Ordway^2,5^,
Marielle N. Smith^6^,
Tyeen Taylor^7^,
Lawren Sack^2^,
Kristina J. Anderson-Teixeira^1,3^*

@@

@@

Tree metabolism and growth are shaped by the positioning of their crowns within the vertical gradient.
Tree height, crown volume, and foliage biomass all scale with diameter at breast height (DBH), which in turn is a strong predictor of tree transpiration  [@meinzerWaterTransportTrees2001;@kunertRevisedHydrologicalModel2017; @anderson-teixeiraSizerelatedScalingTree2015] and photosynthesis.
Specifically, increases are linked to increased leaf area and the increasing probability that the crown is in the canopy [@muller-landauTestingMetabolicEcology2006], where higher light availability results in higher leaf area-specific photosynthesis (Table `r table_leaf_metabolism`).
The net foliar photosynthate production is allocated among functions including respiration, above-ground woody growth, foliar turnover, root growth and allocation to root-associated microorganisms, reproduction, defense, and storage of non-structural carbohydrates (NSCs). 
Among these, the process about which we know the most is woody aboveground growth, which consumes only a modest fraction of total photosynthate [~1/6 on the ecosystem level, @anderson-teixeiraCarbonCyclingMature2021] but is disproportionately important to long-term forest dynamics and carbon cycling in that it builds up woody tissues with a long residence time in the ecosystem [@russellResidenceTimesDecay2014].
Radial stem growth may increase or decrease over time as trees grow in DBH depending on the light environment [@anderson-teixeiraJointEffectsClimate2022].
In open forests and for open-growth individual trees, growth rate declines with DBH [@muller-landauTestingMetabolicEcology2006; @anderson-teixeiraJointEffectsClimate2022], whereas growth rate consistently increases with DBH for trees established in the understory of a closed-canopy forest [@muller-landauTestingMetabolicEcology2006; @anderson-teixeiraSizerelatedScalingTree2015]. 
This points to a dominant role of vertical profiles in microclimate in shaping tree growth rates within forests.

>>>>

**Authors:** 
Nidhi Vinod^1,2^,
Martijn Slot^3^,
Ian McGregor^4^,
Elsa M. Ordway^2,5^,
Marielle N. Smith^6^,
Tyeen Taylor^7^,
Lawren Sack^2^,
Tom Buckly^8^,
Kristina J. Anderson-Teixeira^1,3^*

@@

8. Department of Plant Sciences; University of California; Davis, CA, 95616, USA

@@

Tree metabolism and growth are shaped by the positioning of their crowns within the vertical gradient.
Tree height, crown volume, and foliage biomass all scale with diameter at breast height (DBH), which in turn is a strong predictor of tree transpiration  [@meinzerWaterTransportTrees2001;@kunertRevisedHydrologicalModel2017; @anderson-teixeiraSizerelatedScalingTree2015] and photosynthesis.
Specifically, increases are linked to increased leaf area and the increasing probability that the crown is in the canopy [@muller-landauTestingMetabolicEcology2006], where higher light availability results in higher leaf area-specific photosynthesis (Table `r table_leaf_metabolism`).
The net foliar photosynthate production is allocated among functions including respiration, above-ground woody growth, foliar turnover, root growth and allocation to root-associated microorganisms, reproduction, defense, and storage of non-structural carbohydrates (NSCs). 
Among these, the process about which we know the most is woody aboveground growth, which consumes only a modest fraction of total photosynthate [~1/6 on the ecosystem level, @anderson-teixeiraCarbonCyclingMature2021] but is disproportionately important to long-term forest dynamics and carbon cycling in that it builds up woody tissues with a long residence time in the ecosystem [@russellResidenceTimesDecay2014].
Radial stem growth may increase or decrease over time as trees grow in DBH depending on the light environment [@anderson-teixeiraJointEffectsClimate2022].
In open forests and for open-growth individual trees, growth rate declines with DBH [@muller-landauTestingMetabolicEcology2006; @anderson-teixeiraJointEffectsClimate2022], whereas growth rate consistently increases with DBH for trees established in the understory of a closed-canopy forest [@muller-landauTestingMetabolicEcology2006; @anderson-teixeiraSizerelatedScalingTree2015]. 
This points to a dominant role of vertical profiles in the biophysical environment (Fig. `r fig_NEON_vertical`) in shaping tree growth rates within forests.
